---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.


1. Inspect the first few lines of the data set. What variables are there? Of what type are the variables? What does each variable mean? What do we expect their data range to be?

```{r}
library(classdata)
library(ggplot2)
library(dplyr)
df = ames %>% filter(YearBuilt!=0) %>% filter(`LotArea(sf)`!=0) %>% filter(Acres!=0) %>% filter(`TotalLivingArea (sf)`!=0) %>% filter(`Sale Price`!=0) %>% filter(Bedrooms!=0)
ames = df
summary(ames)
head(ames)
str(ames)
```
The variables of the data set are: Parcel ID, Address, Style, Occupancy, Sale Date, Sale Price, Multi Sale, YearBuilt, Acres, TotalLivingArea (sf), Bedrooms, FinishedBsmtArea (sf), LotArea(sf), AC, FirePlace, and Neighborhood. That makes 17 variables, and we removed the rows with strange values of 0 for certain variables. Parcel ID is of type chr, and is a 10 digit number providing a unique ID to each residence. Address is of type chr, and is a string representing the location of the residence. Style is of type Factor, and represents the general size of the residence in stories. Occupancy is of type Factor, and represents the type of residence, such as Townhouse or Condominium. Sale date is of type date, represents the date the residence was sold, and go from 2017-07-03 to 2022-08-31. Sale price is of type num, represents the USD cost of the residence, and ranges from $1 to $20,500,000. Multi Sale is of type chr, and represents if the residence sale occurred with others. Most of these values are NA. YearBuilt is of type num, represents the year the residence was built, and ranges from 1880 to 2021.


2. Is there a variable of special interest or focus?

  For this model, We are focusing on Sale Price.
 
3. Start the exploration with the main variable. What is the range of the variable? Draw a histogram for a numeric variable or a bar chart, if the variable is categorical. What is the general pattern? Is there anything odd? Follow up on any oddities in step 4.


```{r }
min(ames$'Sale Price')
max(ames$'Sale Price')
options(scipen = 999) # Option to prevent scientific notation
ggplot(ames, aes(x = `Sale Price`)) +
  geom_histogram(bins = 30) +
  labs(title = "Histogram of Sale Price",
       x = "Sale Price",
       y = "Frequency")
options(scipen = 0) # Resetting option to default
```

The main variable "Sale Price" ranges from 1 to 20.5 million dollars. When looking at a histogram of "Sale Price", we see that the overwhelming majority of the data is contained within the first three bins, and the rest of the data is contained within two bins on the higher side of the range of "Sale Price". It is odd that there are so many possible values for "Sale Price" that don't have any observations to match, leaving us with a lot of empty space in the plot.

4. Pick a variable that might be related to the main variable. What is the range of that variable? Plot/describe the pattern. What is it's relationship to the main variable? Plot a scatterplot, boxplot, or faceted barcharts (depending on the types of variables involved). Describe overall pattern, does this variable describe any oddities discovered in 3? Identify/follow-up on any oddities.

Niraj: I am choosing the YearBulit variable as my secondary variable to explore with sales variable. 

```{r, include=FALSE}
library(tidyverse)
max(ames$`YearBuild (sf)`, na.rm = TRUE)
min(ames$`YearBuild (sf)`, na.rm = TRUE)
ggplot(data = ames, aes(x = 'YearBuild (sf)')) + geom_bar()

```
Vinayak: I am doing the TotalLivingArea (sf)
The range of this variable is:
```{r, include=FALSE}
library(tidyverse)
max(ames$`TotalLivingArea (sf)`, na.rm = TRUE)
min(ames$`TotalLivingArea (sf)`, na.rm = TRUE)

ggplot(data = ames, aes(x = 'TotalLivingArea (sf)')) + geom_bar()

```
Relationship to the main variable:

