---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

**TL;DR** Sales price is the most important variable, and follows a linear relationship with Total Living Area. Acres and Sales Price also follow a linear relationship, albeit with a much lower strength. Occupancy does have an effect on Sale Price, given the differing boxplots for each level.

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.


1. Inspect the first few lines of the data set. What variables are there? Of what type are the variables? What does each variable mean? What do we expect their data range to be?

```{r}
library(classdata)
library(ggplot2)
library(dplyr)
df = ames %>% filter(YearBuilt!=0) %>% filter(`LotArea(sf)`!=0) %>% filter(Acres!=0) %>% filter(`TotalLivingArea (sf)`!=0) %>% filter(`Sale Price`!=0) %>% filter(Bedrooms!=0) %>% filter(Acres!=0) %>% filter(`TotalLivingArea (sf)`!=0) %>% filter(`Bedrooms` != 0) %>% filter(`FinishedBsmtArea (sf)`!=0) %>% filter(AC != 0) %>% filter(FirePlace != 0) %>% filter(Neighborhood != 0)
ames = df
summary(ames)
head(ames)
str(ames)

class(df$Acres)
min(df$Acres)
max(df$Acres)

class(df$`TotalLivingArea (sf)`)
min(df$`TotalLivingArea (sf)`)
max(df$`TotalLivingArea (sf)`)

class(df$Bedrooms)
min(df$Bedrooms)
max(df$Bedrooms)

class(df$`FinishedBsmtArea (sf)`)
min(df$`FinishedBsmtArea (sf)`)
max(df$`FinishedBsmtArea (sf)`)

class(df$`LotArea(sf)`)
min(df$`LotArea(sf)`)
max(df$`LotArea(sf)`)

class(df$AC)
min(df$AC)
max(df$AC)

class(df$FirePlace)
min(df$FirePlace)
max(df$FirePlace)

class(df$Neighborhood)


```
The variables of the data set are: Parcel ID, Address, Style, Occupancy, Sale Date, Sale Price, Multi Sale, YearBuilt, Acres, TotalLivingArea (sf), Bedrooms, FinishedBsmtArea (sf), LotArea(sf), AC, FirePlace, and Neighborhood. That makes 17 variables, and we removed the rows with strange values of 0 for certain variables. Parcel ID is of type chr, and is a 10 digit number providing a unique ID to each residence. Address is of type chr, and is a string representing the location of the residence. Style is of type Factor, and represents the general size of the residence in stories. Occupancy is of type Factor, and represents the type of residence, such as Townhouse or Condominium. Sale date is of type date, represents the date the residence was sold, and go from 2017-07-03 to 2022-08-31. Sale price is of type num, represents the USD cost of the residence, and ranges from $1 to $20,500,000. Multi Sale is of type chr, and represents if the residence sale occurred with others. Most of these values are NA. YearBuilt is of type num, represents the year the residence was built, and ranges from 1880 to 2021.

Acres represents the acerage of the properties, is of type 'double' and ranges from 0.001 to 4.65.
TotalLivingArea represents the living area of the properties, is of type 'double' and ranges from 3 to 6007 sf.
Bedrooms represents the number of bedrooms per property, is of type 'double', and ranges from 1 to 10.
FinishedBsmtArea represents the basement area of each property, is of type 'double', and ranges from 16 to 2537 sf.
LotArea represents the lot area of each property, is of type 'double', and ranges from 1037 to 202554.
AC represents whether or not there is an AC in the property, is of type 'character', and is either No or Yes.
Fireplace represents whether or not there is a fireplace on the property, is of type 'character', and is either Yes or No.
Neighborhood represents the neighborhood of the property and is of type 'factor'.










2. Is there a variable of special interest or focus?

  For this model, We are focusing on Sale Price.
 
3. Start the exploration with the main variable. What is the range of the variable? Draw a histogram for a numeric variable or a bar chart, if the variable is categorical. What is the general pattern? Is there anything odd? Follow up on any oddities in step 4.


```{r }
min(ames$'Sale Price')
max(ames$'Sale Price')
options(scipen = 999) # Option to prevent scientific notation
ggplot(ames, aes(x = `Sale Price`)) +
  geom_histogram(bins = 30) +
  labs(title = "Histogram of Sale Price",
       x = "Sale Price",
       y = "Frequency")
options(scipen = 0) # Resetting option to default
```

The main variable "Sale Price" ranges from 1 dollar to 1.2 million dollars. When looking at a histogram of "Sale Price", we see that the data has a unimodal shape with a slight right skew. There are some outliers towards the higher end of the range, suggesting that our data has values that could potentially alter further analysis. Besides those outliers, most of the data seems to be distributed close to the mean, which is around 250 thousand dollars.

4. Pick a variable that might be related to the main variable. What is the range of that variable? Plot/describe the pattern. What is it's relationship to the main variable? Plot a scatterplot, boxplot, or faceted barcharts (depending on the types of variables involved). Describe overall pattern, does this variable describe any oddities discovered in 3? Identify/follow-up on any oddities.

Niraj: I am choosing the YearBulit variable as my secondary variable to explore with sales variable. 

Range of the YearBuild: 

```{r, include=TRUE}

# library(tidyverse)
# library(ggplot2)

range(as.numeric(as.character(ames$YearBuilt)), na.rm = TRUE)

ggplot(data = ames, aes(x = factor(YearBuilt), weight = `Sale Price`))  + geom_bar() + labs(x = 'Year Built') + coord_flip()

```
  
Relationship with the Main Variable( sale Price):

```{r, include=TRUE}

ggplot(data = ames, aes(x = YearBuilt, y = log(`Sale Price`), alpha = 0.2)) +
  geom_point()

```


Conclusion : By observing the Scatter plot, We came to know that there is some relationship between two variables sale price and YearBuilt. As the yearBuilt goes on increasing, the sale price slightly increasing showing the positive association.  There are  oddities between Sale Price and YearBuilt. It means at some years the sale price exponentially increased and decreased compare to other YearBuilt. 

Vinayak: I am doing the TotalLivingArea (sf)
The range of this variable is:
```{r}
max(ames$`TotalLivingArea (sf)`, na.rm = TRUE)
min(ames$`TotalLivingArea (sf)`, na.rm = TRUE)

ggplot(ames, aes(x = `TotalLivingArea (sf)`)) + geom_bar() + coord_flip()

```

Relationship to the main variable:
```{r}
ggplot(data = ames, aes(x = `TotalLivingArea (sf)`, y = `Sale Price`, alpha = 0.2)) + geom_point()

```

Sale Price increases as Total Living Area increases, we see that there are few properties with a living area above 4000 sf, and there is only one property above $1000000. Most properties an area of fewer than 2000 sf. Sale Price and TotalLivingArea have a mostly linear relationship. There aren't any oddities between Sale Price and TotalLivingArea.

Zane: I am choosing the Acres variable.

```{r}
max(ames$Acres, na.rm=TRUE)
min(ames$Acres, na.rm=TRUE)
mean(ames$Acres, na.rm=TRUE)
ggplot(ames, aes(x = Acres)) + 
  geom_histogram(binwidth=0.1) +
  ggtitle("binwidth = 0.1")
```

The distribution is heavily skewed right, with a minimum of 0.024 acres and a maximum of 4.65 acres. The mean acreage is 0.266.

```{r}
ggplot(ames, aes(x = Acres, y = `Sale Price`, alpha = 0.2)) +
  geom_point()
```

The majority of the residences in the scatterplot are clustered with low acreage and low to medium price. There is no clear correlation due to the extreme density of points. However, we can tell that there is a positive correlation between acreage and sale price.There are no obvious oddities between acreage and sale price.


Ryan: I will look at Occupancy.
```{r }
summary(ames$Occupancy)
```

There are 5 levels to this categorical variable: Condominium, Townhouse, Two-Family Duplex, Single-Family/Owner Occupied, Two-Family Conversion.

```{r }
ggplot(data = ames, aes(x = Occupancy)) +
  geom_bar()
```

The vast majority of observations fall under Single-Family/Owner Occupied; there are a fewer but still a good amount of Townhouses; Two-Family Conversion and Two-Family Duplex have similarly small frequencies; Condominium is clearly the least represented group.

```{r }
options(scipen = 999)
ggplot(data = ames) +
  geom_boxplot(mapping = aes(x = `Sale Price`, y = Occupancy))
options(scipen = 0)
```

In order from least to greatest median Sale Price we have: Condominium, Two-Family Conversion, Two-Family Duplex and Townhouse roughly equivalent, and Single-Family/Owner Occupied. Apart from Condominium, each level of Occupancy has outliers on the high end. Single-Family/Owner Occupied is the biggest culprit of this, with a significant number of these high-end outliers. This are likely to be related to the outliers seen in the original histogram of Sale Price. Because each of the boxplots are different, we can say that Occupancy does have an effect on Sale Price.
